### 2024-08-20 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Added ICC and updated graph

---
title: "Multiple regression: TRELIS-Q WP5 data"
author: "Niklas Karlsen"
date: "2024-08-13"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/bookchapter.TRELIS-Q-WP5.html')) })
---

TODO

* endre navn til paper 2

Changelog:

* 13.8.2024: added check of ICC
* 16.5.2024: added plot for NFSUN2024

# Initialize

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #,results="asis")

Sys.setenv(LANG = "no_NB.UTF-8")            # for norwegian æ,ø og å. Kilde:
Sys.setlocale("LC_ALL","Norwegian")
# https://stackoverflow.com/questions/16347731/how-to-change-the-locale-of-r

plot_to_file = 1
```

## Load libraries

```{r load_libraries, echo=FALSE, warning=FALSE, message=FALSE}
#library(pander)                             # prettyprint for tables
library(plyr)
library(ggplot2)
library(viridis)

library(mice)                               # for md.pattern (missing)
library(psych)                              # for ICC, SD

# cluster-variable
# kilde: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/
library(estimatr)                           # for cluster corrected SE
#library(sandwich)
#library(lmtest)

library(nlme)                               # for ICC - variance HLM
```

## Help functions

```{r}
# describe data (mean, sd, skew, curtosis)
describe_data <- function (datalist, QUIET=1) {
  im <- mean(datalist, na.rm=TRUE)
  isd <- sd(datalist, na.rm=TRUE)
  isk <- psych::skew(datalist, na.rm=TRUE) # from psych
  ik <- kurtosi(datalist, na.rm=TRUE) # from psych
  mis <- sum(is.na(datalist)) # missing
  ir <- max(datalist, na.rm=TRUE) -          
    min(datalist, na.rm=TRUE) # range

  # source: https://stackoverflow.com/questions/34351598/getting-name-of-variable-in-r
  if (QUIET != 1) {
    print(sprintf("%7s = %s", "Item", deparse(substitute(datalist))))
    print(sprintf("%7s = %.3f", "mean", im))
    print(sprintf("%7s = %.3f", "std.dev", isd))
    print(sprintf("%7s = %.3f", "skew", isk))
    print(sprintf("%7s = %.3f", "kurt", ik))
    print(sprintf("%7s = %.3f", "range", ir))
    print(sprintf("%7s = %i", "missing", mis))

    if (isk > 2 | ik > 7) {
      print("                            Skew or curtosis TOO LARGE.")
    }
  }

  # outliers
  lowerq = quantile(datalist, na.rm=TRUE)[2]
  upperq = quantile(datalist, na.rm=TRUE)[4]
  iqr = upperq - lowerq #Or use IQR(df)
  # gang med 3 for extreme outliers, 1.5 for mild outliers

  mild.threshold.upper = (iqr * 3) + upperq
  mild.threshold.lower = lowerq - (iqr * 3)

  # print outliers
  # kilde: https://stackoverflow.com/questions/19353116/how-do-i-print-values-in-a-list-that-are-greater-than-a-certain-number-along-wit#19353192
  outliers_upper <- which(datalist>mild.threshold.upper)
  outliersu <- datalist[outliers_upper]
  outliers_lower <- which(datalist<mild.threshold.lower)
  outliersl <- datalist[outliers_lower]

  if (length(outliersu) > 0 | length(outliersl) > 0) {
    print("outliers upper =")
    print(outliersu)
    print("outliers lower =")
    print(outliersl)
  }

  #hist (datalist, main=i) # histogram
  #readline(prompt = "Press enter")

  return(c("mean"=round(im,3), "skew"=round(isk,3), "kurtosi"=round(ik,3)))
           #round(outliersu,3), round(outliersl,3)))
}

###
# helper functions from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
###

## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  #library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column  ; from plyr package
  datac <- plyr::rename(datac, c("mean" = measurevar))
  #datac <- rename(datac, measurevar = "mean") # oppdatert til ny versjon av dplyr

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}

## Norms the data within specified grupper in a data frame; it normalizes each
## subject (identified by idvar) so that they have the same mean, within each group
## specified by betweenvars.
##   data: a data frame.
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   na.rm: a boolean that indicates whether to ignore NA's
normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
                           na.rm=FALSE, .drop=TRUE) {
  #library(plyr)

  # Measure var on left, idvar + between vars on right of formula.
  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
                         .fun = function(xx, col, na.rm) {
                           c(subjMean = mean(xx[,col], na.rm=na.rm))
                         },
                         measurevar,
                         na.rm
  )

  # Put the subject means with original data
  data <- merge(data, data.subjMean)

  # Get the normalized data in a new column
  measureNormedVar <- paste(measurevar, "_norm", sep="")
  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
    mean(data[,measurevar], na.rm=na.rm)

  # Remove this subject mean column
  data$subjMean <- NULL

  return(data)
}

```

## Define constructs

```{r load_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_PST_list    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

# Data management

## Load data

```{r load_data_cfa}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_2023.Rdata")
```

## Prepare data

Flytt origo

```{r}
data_trelis_q_ren[,unlist(TPACK_PST_list)] <- data_trelis_q_ren[,unlist(TPACK_PST_list)] - 3 
```

OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett

```{r}
# identify NA as studiested 5(?) 
data_trelis_q_ren[data_trelis_q_ren$studiested == "Nord universitet","Trinn"] <- "Trinn 5-10"
```

## Check for missing

```{r}
# covariates
md.pattern(data_trelis_q_ren[,c("studiested","Gender", "Trinn", "semester")])

# items / likert data
md.pattern(data_trelis_q_ren[, unlist(TPACK_PST_list)])
```

```{r fix_missing, eval=T}
# remove NMAR + NTNU (19)
data_clean <- data_trelis_q_ren[-c(19,34,44,45,46,95,102),]

# jeg skjuler Annet, siden det bare er en person; alternativt NA?
data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
#trinn_r[is.na(trinn_r)] <- 1

# jeg skulet Annet / ønsker ikke å oppgi, siden det bare er to personer
# skjuler som Kvinne, siden det er flest av de
data_clean$Gender[data_clean$Gender == "Annet/ønsker ikke å oppgi"] <- NA # "Kvinne"
#gender_r[is.na(gender_r)] <- 1

# merge 2. and 3. year (Naturfag 1)
#data_clean$semester[data_clean$semester == "2. år"] <- "3. år"
```

# Descriptive statistics

```{r}
scores_v23 <- scoreItems(TPACK_PST_list, data_clean)

describe(scores_v23$scores)

# constructs for multiple regression
TK_v23    <- scores_v23$scores[,"TK"]
PCK_v23   <- scores_v23$scores[,"PCK"]
TPACK_v23 <- scores_v23$scores[,"TPACK"]

# add constructs to data.frame
data_clean$TK    <- TK_v23
data_clean$PCK   <- PCK_v23
data_clean$TPACK <- TPACK_v23
```

## Bargraphs: comparing institutes

```{r}
# group institutes into three: low, medium, high
data_clean$studiested2 <- factor(data_clean$studiested, 
                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
                                 label=c("A", "A", "B", "C", "C"))

dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested2")
dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested2")
dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested2")

# rename; kilde: https://www.statology.org/rename-single-column-in-r/
colnames(dc1)[colnames(dc1) == "TK"] <- "var"
colnames(dc2)[colnames(dc2) == "PCK"] <- "var"
colnames(dc3)[colnames(dc3) == "TPACK"] <- "var"

# merge data
dc <- rbind(dc1,dc2)
dc <- rbind(dc,dc3)

# add column med navn
dc["TPACK"] <- c("TK","TK","TK","PCK","PCK","PCK","TPACK","TPACK","TPACK")
```

```{r}
stolpediagram <- ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("PCK, TK og TPACK fordelt på studiested") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("verdi") +
  xlab("studiested gruppe") +
  theme_bw() +
  facet_wrap(~TPACK)

print(stolpediagram) # vis plot
```

```{r}
# save plot
if (plot_to_file) {
  ggsave("../output/TPACK_stolpediagram.png") #,height=8,width=15)
  }

```

For NFSUN2024:

```{r}
# anonymise institutes
data_clean$studiested3 <- factor(data_clean$studiested, 
                                 level=c("HVL","UiT","Nord universitet","Høgskolen i Innlandet","OsloMet"),
                                 label=c("I1", "I2", "I3", "I4", "I5"))

dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested3")
dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested3")
dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested3")

# rename; kilde: https://www.statology.org/rename-single-column-in-r/
colnames(dc1)[colnames(dc1) == "TK"] <- "var"
colnames(dc2)[colnames(dc2) == "PCK"] <- "var"
colnames(dc3)[colnames(dc3) == "TPACK"] <- "var"

# merge data
dc <- rbind(dc1,dc2)
dc <- rbind(dc,dc3)

# add column med navn
dc["TPACK"] <- c("TK","TK","TK","TK","TK",
                 "PCK","PCK","PCK","PCK","PCK",
                 "TPACK","TPACK","TPACK","TPACK","TPACK")
```

```{r}
stolpediagram <- ggplot(dc, aes(x=studiested3, y=var, fill=studiested3)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("Mean and standard deviation for PCK, TK and TPACK") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("Value") +
  xlab("Institute") +
  theme_bw() +
  facet_wrap(~TPACK)

print(stolpediagram) # vis plot
```

```{r}
# save plot
if (plot_to_file) {
  ggsave("../output/NFSUN2024_stolpediagram.png",dpi=300) #height=8,width=15)
  }

```

```{r eval=F}
ggplot(dc, aes(x=TPACK, y=var, fill=studiested2)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("PCK, TK og TPACK fordelt på studiested") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("verdi") +
  xlab("") +
  theme_bw() +
  facet_wrap(~studiested2)
```

## Check for interactions

### TK Level 1

```{r TK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

Different signs on slope, a clear indication of interaction effects on TK from gender and target grade level.

```{r TK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r TK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### TK Level 2

```{r TK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r TK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### PCK Level 1

```{r PCK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

Different signs on slope, a clear indication of interaction effects on PCK from gender and target grade level.

```{r PCK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r PCK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### PCK Level 2

```{r PCK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r PCK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r PCK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### TPACK Level 1

```{r TPACK_gender_trinn}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Target grade level")
```

Not so clear interaction effect, similar slopes? They do cross eventually.

```{r TPACK_gender_semester}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Study year")
```

```{r TPACK_trinn_semester}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Study year")
```

### TPACK Level 2

```{r TPACK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r TPACK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TPACK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

Not so clear interaction on TPACK from gender and institute.

# Regression analysis

## Check assumptions for regression

### ICC

ICC: Christophersen (2018, p. 109-111)

```{r ICC}
# variance decomposition
null.mod <- lme(TPACK ~ 1, random =~ 1|studiested, data = data_clean)
vcn <- VarCorr(null.mod)
vcn
```

The variance of indviduals is given by the residual. The variance of institutions is givene by the intercept. If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.

### Normality

```{r normality}
# covariates; 80/20 distribution
# ref http://stackoverflow.com/questions/21639392/ddg#21639445
barplot(prop.table(table(data_clean$Gender)))
barplot(prop.table(table(data_clean$Trinn)))
barplot(prop.table(table(data_clean$semester)))
#hist(oslomet_r)
#hist(innlandet_r)
#hist(HVL_r)
#hist(UIT_r)

# variable
hist(data_clean$TK)
hist(data_clean$PCK)
hist(data_clean$TPACK)
```

```{r}
describe_data(data_clean$TK)
describe_data(data_clean$PCK)
describe_data(data_clean$TPACK)
```

### Linearity

```{r scatterplot_linearity, eval=F}
plot(data_clean$Gender,        data_clean$TPACK)
plot(data_clean$Trinn,         data_clean$TPACK)
plot(data_clean$semester,      data_clean$TPACK)
#plot(data_clean$oslomet_r,         data_clean$TPACK_v23)
#plot(data_clean$innlandet_r,       data_clean$TPACK_v23)
#plot(data_clean$HVL_r,             data_clean$TPACK_v23)
#plot(data_clean$UIT_r,             data_clean$TPACK_v23)

plot(data_clean$TK,            data_clean$TPACK)
#plot(data_clean$TK[oslomet_r],data_clean$TPACK[oslomet_r]) # gr.A
#plot(data_clean$TK[innlandet_r],data_clean$TPACK[innlandet_r]) # gr.B
#plot(data_clean$TK[HVL_r],data_clean$TPACK[HVL_r]) # gr.A
#plot(data_clean$TK[UIT_r],data_clean$TPACK[UIT_r]) # gr.B

#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(studiested))
#scatterplot(TPACK_regdata, c("TK", "TPACK"), factor(semester))

plot(data_clean$PCK,           data_clean$TPACK)
```

```{r multicollinearity, eval=F}
#cor(TK_v23, studiested_r)
#cor(PCK_v23, studiested_r)
cor(data_clean$PCK,   data_clean$semester)
#cor(TPACK_v23, studiested_r)
cor(data_clean$TPACK, data_clean$Gender)

cor(data_clean$TK,    data_clean$Gender)
cor(data_clean$TK,    data_clean$Trinn)

cor(data_clean$TK,    data_clean$PCK)
cor(data_clean$TK,    data_clean$TPACK)
cor(data_clean$PCK,   data_clean$TPACK)
```

TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)

## Hierarchical RA for TK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r TKlevel2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r TKlevel2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
# interaction
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for PCK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r PCKlevel2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCKlevel2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for TPACK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r level2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)

#main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r level2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)

fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### PCK

```{r PCK}
# PCK
m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK), 
                clusters = factor(studiested), data=data_clean)
summary(m5)

# main
fstat <- m5$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m5$adj.r.squared / m4$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### TK

```{r TK}
# TK
m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK) + scale(TK), 
                clusters = factor(studiested), data=data_clean)
summary(m6)

fstat <- m6$adj.r.squared / m5$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```