### 2024-08-22 Rmd/thesis-reanalyses.Rmd

---
title: "Reanalyses done for the synopsis"
author: "Niklas Karlsen"
date: "2024-08-21"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/reanalyses-thesis.html')) })
---

```{r}
# load libraries
library(nlme)
library(psych)
library(dplyr)
library(pander)
```

# Introduction

Here I present the reanalyses work that was done as part of writing the synopsis. 
# Methods

## Samples

Load data

```{r load_data_survey}
# Rdata-file contains reduced / prepared datasets
load(file="../../data/processed_data/tpack-TRELIS-Q_2023.Rdata")
```

```{r fix_missing, eval=T}
# remove NMAR + NTNU (19)
data_clean <- data_trelis_q_ren[-c(19,34,44,45,46,95,102),] 
```

## Instruments

### Define constructs

```{r load_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_PST_list    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

## Calculate mean values

```{r}
scores_v23 <- scoreItems(TPACK_PST_list, data_clean)

describe(scores_v23$scores)

# constructs for multiple regression
TK_v23    <- scores_v23$scores[,"TK"]
PCK_v23   <- scores_v23$scores[,"PCK"]
TPACK_v23 <- scores_v23$scores[,"TPACK"]

# add constructs to data.frame
data_clean$TK    <- TK_v23
data_clean$PCK   <- PCK_v23
data_clean$TPACK <- TPACK_v23
```

# Examine correlation between TK and TPACK

## Survey

level 2-analysis

```{r}
#lme(TPACK ~ TK, )

hlm <- lme(TPACK ~ TK, random=~1|studiested, method="REML", data=data_clean)
summary(hlm)
```

The correlation between the intercept (value of TPACK at each teacher education institute (TEI)) and TK is negative, indicating that the higher TPACK is at a TEI, the lower the correlation with TK (i.e., the slope) becomes (Brown, 2021). This could be a ceiling effect, but since TPACK is relatively low, it probably isn't?

```{r}
# per institute and year of study
mTK <- aggregate(x=data_clean$TK,
          by=list(data_clean$studiested,data_clean$semester),
          FUN=mean)

mTPACK <- aggregate(x=data_clean$TPACK,
          by=list(data_clean$studiested,data_clean$semester),
          FUN=mean)

colnames(mTK) <- c("studiested", "semester", "mTK")
colnames(mTPACK) <- c("studiested", "semester", "mTPACK")

plot(mTK[,3], mTPACK[,3]) # x contains mean value
```

Plot correlation (TK/TPACK) vs TPACK

```{r}
# ref: https://stats.stackexchange.com/questions/4040/r-compute-correlation-by-group [21.8.2024]
#r_TK_TPACK <- by(data_clean[,c("TK", "TPACK")], data_clean$studiested, function(x) {cor(x$TK, x$TPACK)})

# ref: https://www.geeksforgeeks.org/how-to-calculate-correlation-by-group-in-r/ [21.8.2024]
r_TK_TPACK <- data_clean %>%
  group_by(studiested, semester) %>%
  summarise(correlation = cor(TK, TPACK))

# ref: https://www.statology.org/r-merge-multiple-columns/ [21.8.2024]
cor_TPACK_table <- merge(r_TK_TPACK, mTK, by.x=c('studiested', 'semester'), by.y=c('studiested', 'semester'))

cor_TPACK_table <- merge(cor_TPACK_table, mTPACK, by.x=c('studiested', 'semester'), by.y=c('studiested', 'semester'))

pander(cor_TPACK_table)

plot(cor_TPACK_table$correlation, cor_TPACK_table$mTPACK)
```

```{r}
# remove outlier correlation, which is also a indication of TK and TPACK being partially unrelated
linmod <- lm(correlation ~ mTPACK,
             data=cor_TPACK_table[cor_TPACK_table$correlation>0.5,])

summary(linmod)
```

## Learning sequence

Correlation between pre-values compared to correlation between post-values.