### 2024-09-04 Rmd/Paper2.Rmd
Several corrections

---
title:        "Paper 2: survey analysis"
author:       "Niklas Karlsen"
date:         "2024-09-04"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/paper2.html')) })
---

# Introduction

This document presents the quantitative analyses done in paper 2. Descriptive and inferential analyses are done. The inferences are based on multiple regression analyses.

## Load libraries

```{r load_libraries}

### prepare raw data
library(readxl)                             # read .xlsx-files. Kilde: https://readxl.tidyverse.org/

### presentation of results
#library(pander)                             # prettyprint for tables

# NB! Collision between plyr and dplyr!
library(plyr)
library(dplyr)                              # rename columns

library(ggplot2)
library(viridis)

### analyses
library(mice)                               # for md.pattern (missing)
library(psych)                              # for ICC, SD

# cluster-variable
# kilde: https://www.r-bloggers.com/2021/05/clustered-standard-errors-with-r/
library(estimatr)                           # for cluster corrected SE
#library(sandwich)
#library(lmtest)
library(nlme)                               # for ICC - variance HLM
```

## Define constructs

The constructs are based on @karlsen_assessing_2024.

```{r load_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

# Method

The following approach is used in this document. First, the raw data is loaded and prepared for analyses. Second, descriptive statistics are created. Third, inferential statistics (i.e., regression analyses) are created. Finally, some comments are provided on the results.

## Data management

For the survey, the complete TRELIS-Q dataset was used.

### Load data

```{r}
folder_trelis_q <- "../data/rawdata/"
trelis_q <- "trelis-q-datafil-endelig-juni2023.xlsx"

path_trelis_q <- paste(folder_trelis_q, trelis_q, sep='')

data_trelis_q <- read_excel(path_trelis_q, sheet=4)
```

### Data cleaning

```{r}
# only keep relevant columns
data_trelis_q <- data_trelis_q[,51:74]
```

```{r}
# keep some columns; covariates
# 68. Hvilket studiested studerer du ved?	
# 69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?	
# 70. Hvilket studieår er du i?	
# 71. Er du:

# likert-spørsmål: 48 - 66
# åpne spørsmål: 67

data_trelis_q <- rename(data_trelis_q,
                            "Trinn" =
                            "69. Går du på grunnskolelærerutdanning for trinn 1-7 eller trinn 5-10?")
data_trelis_q <- rename(data_trelis_q,
                            "Gender" =
                            "71. Er du:")
data_trelis_q <- rename(data_trelis_q,
                            "studiested" =
                            "68. Hvilket studiested studerer du ved?")
data_trelis_q <- rename(data_trelis_q,
                            "semester" =
                            "70. Hvilket studieår er du i?")

# change from year to semester
#data_trelis_q$semester <- rep(7, 52)  # 4 år =  7. semester (høst)

# reorder some columns
data_trelis_q <- data_trelis_q %>% relocate(Trinn, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
data_trelis_q <- data_trelis_q %>% relocate(Gender, .before = "48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.")
data_trelis_q <- data_trelis_q %>% relocate(studiested, .before = "Trinn")
data_trelis_q <- data_trelis_q %>% relocate(semester, .before = "Trinn")
```

Rename items to TPACK-lingo. (introduce bias in interpretation? Should it be named something neutral instead?)

```{r}
# rename columns: (bruker dplyr)
# kilde: https://www.sharpsightlabs.com/blog/rename-columns-in-r/
data_trelis_q_ren <- rename(data_trelis_q, 
'CK123' = '48. Jeg forstår sentrale begreper i naturfag innen temaer som energi og materie, jorda og livet på jorda, kropp og helse osv.',
'CK4' = "49. Jeg forstår teknologiske prinsipper og virkemåter som er relevante for naturfag",
'CK5' = '50. Jeg har kunnskaper om naturvitenskapelige praksiser og tenkemåter',
'PCK2' = '51. Jeg vet hvordan jeg kan støtte elevene i å utvikle kritisk tenkning og argumentasjonsferdigheter i naturfag',
'TK2' = "52. Jeg kan sette opp en algoritme når jeg programmerer",
'TK1a' = "53. Jeg kan bryte ned problemer i mindre deler når jeg programmerer",
'TK34' = "54. Jeg har kunnskap om forskjellige programmeringsverktøy, f.eks. Makecode for Micro:bit, Scratch, Python, og lignende",
'TK1b' = "55. Jeg kan oppdage og rette feil, når jeg programmerer.",
'TK567' = "56. Jeg vet hvordan jeg lager kode som inneholder f.eks. variabler, vilkår (if, else) og løkker (for, while)",
'TCK1' = "58. Jeg kan bruke programmering til å utvikle enkle teknologiske systemer som består av deler som virker sammen.",
'TCK4' = '57. Jeg kan bruke programmering til å utforske naturfaglige fenomener',
'PK3' = '59. Jeg kan støtte elevene sosialt i deres læring i ulike fag, f.eks deres nysgjerrighet, kreativitet og samarbeidsevne',
'PK1' = '60. Jeg vet hvordan jeg kan støtte elevenes refleksjon over egen læring i ulike fag',
'PCK5' = '61. Jeg vet hvordan jeg kan vurdere elevenes læring på ulike måter i naturfag',
'TPK1' = '62. Jeg vet hvordan jeg kan bruke programmering på måter som beriker undervisningen i ulike fag',
'TPK4' = '63. Jeg tenker kritisk igjennom hvordan jeg kan bruke programmering i undervisningen i ulike fag',
'TPACK2' = '64. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturfaglige begreper og fenomener',
'TPACK3' = '65. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå naturvitenskapelige praksiser og tenkemåter.',
'TPACK4' = '66. Jeg vet hvordan jeg kan bruke programmering i naturfag på måter som hjelper elevene å forstå hvordan teknologiske systemer fungerer')

# change "5 Stemmer veldig godt" and "1 Stemmer veldig dårlig" to 5 and 1
data_trelis_q_ren[data_trelis_q_ren == '1 Stemmer veldig dårlig'] <- "1"
data_trelis_q_ren[data_trelis_q_ren == '5 Stemmer veldig godt'] <- "5"
```

```{r}
# hva gjør jeg med 9 / VET IKKE? Gjør om til NA?
# tell antall VET IKKE
antall_VETIKKE <- length(which(data_trelis_q_ren == 'VET IKKE'))
print(antall_VETIKKE)
# gjør om til NA; OBS! Ikke gjør disse NA om til tall med impute, pmm senere
data_trelis_q_ren[data_trelis_q_ren == 'VET IKKE'] <- NA # "9"

# convert into numbers
data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)] <-
  apply(
    data_trelis_q_ren[,colnames(data_trelis_q_ren) %in% unlist(TPACK_trelis_q)], 2,
                    function(x) as.numeric(as.character(x)))
```

Hvor mange respondenter er det?

```{r}
nrow(data_trelis_q)
nrow(data_trelis_q_ren)
```

```{r rename_dataset}
data_TRELIS_Q <- data_trelis_q_ren
```

```{r}
# remove open-ended question column
data_TRELIS_Q <- data_TRELIS_Q[,1:23]
```

```{r}
# readxls stores as tibble, not dataframe?
# make tibble into data.frame to treat it similarly to pilot dataset
data_TRELIS_Q <- data.frame(data_TRELIS_Q)
```

### save data

```{r}
save(data_TRELIS_Q,  
     file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
```

```{r}
# clean environment to reset for next step
# ref.: https://www.statology.org/clear-environment-in-r/ [3.9.2024]
rm(list=ls())
```

## Prepare data

```{r load_data_cfa}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
```

```{r reload_constructs}
## TPACK-PST for TRELIS-Q
TK_pos_trelisq    <- c("TK1b","TK2","TK34") 
PCK_pos_trelisq   <- c("CK123","CK5","PCK2","PCK5")
TPACK_pos_trelisq <- c("TPACK2","TPACK4","TCK1","TCK4","TPK1","TPK4")

TPACK_trelis_q    <- list(TK=TK_pos_trelisq,
                     PCK=PCK_pos_trelisq,
                     TPACK=TPACK_pos_trelisq)
```

## Help functions

```{r}
# describe data (mean, sd, skew, curtosis)
describe_data <- function (datalist, QUIET=1) {
  im <- mean(datalist, na.rm=TRUE)
  isd <- sd(datalist, na.rm=TRUE)
  isk <- psych::skew(datalist, na.rm=TRUE) # from psych
  ik <- kurtosi(datalist, na.rm=TRUE) # from psych
  mis <- sum(is.na(datalist)) # missing
  ir <- max(datalist, na.rm=TRUE) -          
    min(datalist, na.rm=TRUE) # range

  # source: https://stackoverflow.com/questions/34351598/getting-name-of-variable-in-r
  if (QUIET != 1) {
    print(sprintf("%7s = %s", "Item", deparse(substitute(datalist))))
    print(sprintf("%7s = %.3f", "mean", im))
    print(sprintf("%7s = %.3f", "std.dev", isd))
    print(sprintf("%7s = %.3f", "skew", isk))
    print(sprintf("%7s = %.3f", "kurt", ik))
    print(sprintf("%7s = %.3f", "range", ir))
    print(sprintf("%7s = %i", "missing", mis))

    if (isk > 2 | ik > 7) {
      print("                            Skew or curtosis TOO LARGE.")
    }
  }

  # outliers
  lowerq = quantile(datalist, na.rm=TRUE)[2]
  upperq = quantile(datalist, na.rm=TRUE)[4]
  iqr = upperq - lowerq #Or use IQR(df)
  # gang med 3 for extreme outliers, 1.5 for mild outliers

  mild.threshold.upper = (iqr * 3) + upperq
  mild.threshold.lower = lowerq - (iqr * 3)

  # print outliers
  # kilde: https://stackoverflow.com/questions/19353116/how-do-i-print-values-in-a-list-that-are-greater-than-a-certain-number-along-wit#19353192
  outliers_upper <- which(datalist>mild.threshold.upper)
  outliersu <- datalist[outliers_upper]
  outliers_lower <- which(datalist<mild.threshold.lower)
  outliersl <- datalist[outliers_lower]

  if (length(outliersu) > 0 | length(outliersl) > 0) {
    print("outliers upper =")
    print(outliersu)
    print("outliers lower =")
    print(outliersl)
  }

  #hist (datalist, main=i) # histogram
  #readline(prompt = "Press enter")

  return(c("mean"=round(im,3), "skew"=round(isk,3), "kurtosi"=round(ik,3)))
           #round(outliersu,3), round(outliersl,3)))
}

###
# helper functions from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
###

## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  #library(plyr)

  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }

  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )

  # Rename the "mean" column  ; from plyr package
  datac <- plyr::rename(datac, c("mean" = measurevar))
  #datac <- rename(datac, measurevar = "mean") # oppdatert til ny versjon av dplyr

  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult

  return(datac)
}

## Norms the data within specified grupper in a data frame; it normalizes each
## subject (identified by idvar) so that they have the same mean, within each group
## specified by betweenvars.
##   data: a data frame.
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   na.rm: a boolean that indicates whether to ignore NA's
normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
                           na.rm=FALSE, .drop=TRUE) {
  #library(plyr)

  # Measure var on left, idvar + between vars on right of formula.
  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
                         .fun = function(xx, col, na.rm) {
                           c(subjMean = mean(xx[,col], na.rm=na.rm))
                         },
                         measurevar,
                         na.rm
  )

  # Put the subject means with original data
  data <- merge(data, data.subjMean)

  # Get the normalized data in a new column
  measureNormedVar <- paste(measurevar, "_norm", sep="")
  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
    mean(data[,measurevar], na.rm=na.rm)

  # Remove this subject mean column
  data$subjMean <- NULL

  return(data)
}

```

Flytt origo

```{r}
data_TRELIS_Q[,unlist(TPACK_trelis_q)] <- data_TRELIS_Q[,unlist(TPACK_trelis_q)] - 3 
```

OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett

```{r}
# identify NA as studiested 5(?) 
data_TRELIS_Q[data_TRELIS_Q$studiested == "Institute3","Trinn"] <- "Trinn 5-10"
```

## Check for missing

The complete TRELIS_Q dataset is almost identical to the trelis_q dataset used in paper 1, except that TRELIS_Q includes all answers from institute2. These were not all collected when paper 1 was written. The addition of these final answers did not add any new NMAR to the dataset.

```{r}
# covariates
md.pattern(data_TRELIS_Q[,c("studiested","Gender", "Trinn", "semester")])

# items / likert data
md.pattern(data_TRELIS_Q[, unlist(TPACK_trelis_q)])
```

```{r fix_missing, eval=T}
# remove NMAR + Institute9 (19)
data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),]

# jeg skjuler Annet, siden det bare er en person; alternativt NA?
data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
#trinn_r[is.na(trinn_r)] <- 1

# jeg skulet Annet / ønsker ikke å oppgi, siden det bare er to personer
# skjuler som Kvinne, siden det er flest av de
data_clean$Gender[data_clean$Gender == "Annet/ønsker ikke å oppgi"] <- NA # "Kvinne"
#gender_r[is.na(gender_r)] <- 1

# merge 2. and 3. year (Naturfag 1)
#data_clean$semester[data_clean$semester == "2. år"] <- "3. år"
```

# Descriptive statistics

```{r}
scores_v23 <- scoreItems(TPACK_trelis_q, data_clean)

describe(scores_v23$scores)

# constructs for multiple regression
TK_v23    <- scores_v23$scores[,"TK"]
PCK_v23   <- scores_v23$scores[,"PCK"]
TPACK_v23 <- scores_v23$scores[,"TPACK"]

# add constructs to data.frame
data_clean$TK    <- TK_v23
data_clean$PCK   <- PCK_v23
data_clean$TPACK <- TPACK_v23
```

## Bargraphs: comparing institutes

```{r}
# group institutes into three: low, medium, high
data_clean$studiested2 <- factor(data_clean$studiested, 
                                 level=c("Institute1","Institute5","Institute3","Institute2","Institute4"),
                                 label=c("A", "A", "B", "C", "C"))

dc1 <- summarySE(data_clean, measurevar="TK", groupvars="studiested2")
dc2 <- summarySE(data_clean, measurevar="PCK", groupvars="studiested2")
dc3 <- summarySE(data_clean, measurevar="TPACK", groupvars="studiested2")

# rename; kilde: https://www.statology.org/rename-single-column-in-r/
colnames(dc1)[colnames(dc1) == "TK"] <- "var"
colnames(dc2)[colnames(dc2) == "PCK"] <- "var"
colnames(dc3)[colnames(dc3) == "TPACK"] <- "var"

# merge data
dc <- rbind(dc1,dc2)
dc <- rbind(dc,dc3)

# add column med navn
dc["TPACK"] <- c("TK","TK","TK","PCK","PCK","PCK","TPACK","TPACK","TPACK")
```

```{r}
stolpediagram <- ggplot(dc, aes(x=studiested2, y=var, fill=studiested2)) + 
  geom_bar(stat="identity", position="dodge") +
  #scale_fill_manual(values=viridis(3)) + 
  geom_errorbar(aes(ymin=var-sd, ymax=var+sd), # se
                  linewidth=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
  ggtitle("PCK, TK og TPACK fordelt på studiested") +
  guides(fill="none") +
  # https://stackoverflow.com/questions/53843221/make-y-axis-start-at-1-instead-of-0-within-ggplot-bar-chart
#  coord_cartesian(ylim = c(1,5)) +
  ylab("verdi") +
  xlab("studiested gruppe") +
  theme_bw() +
  facet_wrap(~TPACK)

print(stolpediagram) # vis plot
```

# Multiple regression analysis

The data is multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2).

## Check assumptions for regression

### ICC

To check whether to treat the data as multilevel, ICC is calculated [@christophersen_introduksjon_2018] (p. 109-111).

```{r ICC}
# variance decomposition
null.mod <- lme(TPACK ~ 1, random =~ 1|studiested, data = data_clean)
vcn <- VarCorr(null.mod)
vcn
```

The variance of indviduals is given by the residual. The variance of institutions is givene by the intercept. If the variance of the intercept is larger than 0.05 of the total variance (intercept + residual), then the institutional variance contributes to explain the outcome.

### Normality

```{r normality}
# covariates; 80/20 distribution
# ref http://stackoverflow.com/questions/21639392/ddg#21639445
barplot(prop.table(table(data_clean$Gender)))
barplot(prop.table(table(data_clean$Trinn)))
barplot(prop.table(table(data_clean$semester)))
#hist(Institute4_r)
#hist(Institute2_r)
#hist(Institute1_r)
#hist(Institute5_r)

# variable
hist(data_clean$TK)
hist(data_clean$PCK)
hist(data_clean$TPACK)
```

```{r}
describe_data(data_clean$TK)
describe_data(data_clean$PCK)
describe_data(data_clean$TPACK)
```

### Linearity

```{r scatterplot_linearity}
# ref: https://stackoverflow.com/questions/29425892/how-do-i-loop-through-column-names-and-make-a-ggplot-scatteplot-for-each-one
for (var1 in names(data_clean)[names(data_clean) %in% c("Gender", "Trinn", "semester", "TK", "PCK")]) {
  xy_plot <- ggplot(data_clean, aes(.data[[var1]],TPACK)) +
        geom_point() +                            # shape=1,alpha=1/5
        geom_jitter() +
        ggtitle(paste("scatterplot",var1,"TPACK"))

  print(xy_plot)
}

#plot(data_clean$Trinn,         data_clean$TPACK)
#plot(data_clean$semester,      data_clean$TPACK)
#plot(data_clean$TK,            data_clean$TPACK)
#plot(data_clean$PCK,           data_clean$TPACK)
```

```{r multicollinearity}
#cor(data_clean$PCK,   data_clean$semester)
#cor(data_clean$TPACK, data_clean$Gender)

#cor(data_clean$TK,    data_clean$Gender)
#cor(data_clean$TK,    data_clean$Trinn)

cor(data_clean$TK,    data_clean$PCK)
cor(data_clean$TK,    data_clean$TPACK)
cor(data_clean$PCK,   data_clean$TPACK)
```

TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)

## Check for interactions

Should I include interactions in the regression analysis?

### TK Level 1

```{r TK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

Different signs on slope, a clear indication of interaction effects on TK from gender and target grade level.

```{r TK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r TK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### TK Level 2

```{r TK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r TK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### PCK Level 1

```{r PCK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

Different signs on slope, a clear indication of interaction effects on PCK from gender and target grade level.

```{r PCK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r PCK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### PCK Level 2

```{r PCK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r PCK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r PCK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### TPACK Level 1

```{r TPACK_gender_trinn}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Target grade level")
```

Not so clear interaction effect, similar slopes? They do cross eventually.

```{r TPACK_gender_semester}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Study year")
```

```{r TPACK_trinn_semester}
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"),
                 trace.label = "Study year")
```

### TPACK Level 2

```{r TPACK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

A clear interaction effect between institute and semester on TPACK.

```{r TPACK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TPACK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

Not so clear interaction on TPACK from gender and institute.

## Hierarchical RA for TK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r TKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r TKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
# interaction
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for PCK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r PCKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for TPACK

Level 1: students  
Level 2: institution

We are using cluster corrected standard errors to correct for clustering on institution.

Add student covariates

### Level 1

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

Calculate F-stat to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r level2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)

#main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r level2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)

fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### PCK

```{r PCK}
# PCK
m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK), 
                clusters = factor(studiested), data=data_clean)
summary(m5)

# main
fstat <- m5$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m5$adj.r.squared / m4$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### TK

```{r TK}
# TK
m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK) + scale(TK), 
                clusters = factor(studiested), data=data_clean)
summary(m6)

fstat <- m6$adj.r.squared / m5$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

# Discussion and conclusion

The results are discussed in @karlsen_assessing_2024.

# References

<!-- citation(package) -->

<div id="refs"></div>

<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html

# (APPENDIX) Appendix {-} -->

# Session info
<!-- # ref.: https://intro2r.com/proj_doc.html -->
<!-- #sessionInfo() -->

```{r}
xfun::session_info()
```