### 2024-10-08 Rmd/Paper4.Rmd
Prepared for publication

---
title: "Paper 4: quasi-experiment"
author: "Niklas Karlsen, OsloMet, Norway, e-mail: niklaska@oslomet.no"
date: "2024-10-08"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), '../output/Paper4.html')) })
---

# Introduction

The purpose of this script is to present the results of the quasi-experiment in paper 4, by presenting the calculations presented in paper 4 for prototype 1 and 2. Three effect sizes are calculated:

* Difference (gain) scores / Hedges' g
* Fixed-effects analysis
* Within-between (hybrid) model

## Load libraries

```{r last_bibliotek}
## bibliotek for tabell-behandling
library(dplyr)
library(tidyr)
library(pander)                             # prettyprint for tables
library(texreg)                               # screenreg (print compare results regresion)

## bibliotek for plotting
library(ggplot2)
library(viridis)

## bibliotek for statistikk
library(psych)                              # for ICC, SD
library(effsize)                            # for cohen's d

### regression libraries
library(lme4)                                 # for MLM
#library(lmerTest)                             # not recommended? Avoid the lmerTest package (https://utstat.toronto.edu/~brunner/workshops/mixed/NormalWithR.pdf)
library(car)                                   # # for p-values i lme4 (anova, test="F"). durbinWatson
library(nlme)                                 # for lme
library(plm)                                  # paneldata analysis

# t-tests
library(WRS2)                                 # Yuen-Welch test ref.: Mair & Wilcox, 2020

# SE calculation
library(pcse)                                 # panel-corrected SE
library(sandwich) # cluster corrected error
library(lmtest) # logLikelihood tests, AIC, BIC ,...
library(aod) # wald.test
library(MuMIn) # AICc

```

## Definer constructs for lettere access i kolonnene

These are the items that were used. The TK-items differ somewhat from the items used in @karlsen_assessing_2024.

```{r load_constructs}
#source("../../scripts/constructs.R")

# TPACK-PST 22
TK_PST22_pos <- c("TK2","TK5","TK6") # ,"TK5","TK6","TCK1"
PCK_PST22_pos <- c("CK1","CK2","CK3","CK5","PCK2","PCK5") # "CK2", "CK5",
TPACK_PST22_pos <- c("TCK1","TCK4","TPK1","TPK4","TPACK2", "TPACK4")

TPACK_PST_emne2_v22 <- list("TK"=TK_PST22_pos,
                            "PCK"=PCK_PST22_pos,
                            "TPACK"=TPACK_PST22_pos)

# TPACK-PST 22 adjusted to 23
TK_PST23_pos <- c("TK2","TK5","TK6") # "TK4","TK5","TK6"
PCK_PST23_pos <- c("CK123", "PCK2")
TPACK_PST23_pos <- c("TPACK2","TCK4","TPK1","TPK4","TPACK4") 
TPACK_PST_emne2_v23 <- list("TK"=TK_PST23_pos,
                            "PCK"=PCK_PST23_pos, 
                            "TPACK"=TPACK_PST23_pos) #, 
```

# Methods

## Samples

### Prepare dataset

Load data

```{r load_data}
# Rdata-file contains reduced / prepared datasets
load(file="../../data/processed_data/tpack-pilot_2022.Rdata")
load(file="../../data/processed_data/tpack-emne2_2023.Rdata")

# fix bug: data_emne2_ren is list not dataframe
data_emne2_ren <- data.frame(data_emne2_ren)
```

Synchronize data structures / naming conventions

```{r create_dataframe_v23}
# [c(1:4,6)] / unlist(TPACK_PST_emne2_v23)
data_emne2_v23 <- data_emne2_ren[,colnames(data_emne2_ren) %in% unlist(TPACK_PST_emne2_v23)]  #- 3

#transfer covariates
data_emne2_v23$Gender   <- c(data_pre1_emne2$Gender, 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$Gender) # data_emne2_ren$Gender
data_emne2_v23$Trinn    <- c(data_pre1_emne2$Trinn, 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$Trinn) # data_emne2_ren$Trinn
data_emne2_v23$praksis  <- c(data_pre1_emne2$'Har du brukt programmering i praksis sammen med skoleelever?', 
                             rep(NA, nrow(data_pre2_emne2)),
                             data_post_emne2$'Har du brukt programmering i praksis sammen med skoleelever?')

data_emne2_v23$full_treatment <- rep(0,nrow(data_emne2_v23))

data_emne2_v23$datasett <- data_emne2_ren$datasett
data_emne2_v23$Subject  <- c(data_pre1_emne2$Subject, 
                             data_pre2_emne2$Subject,
                             data_post_emne2$Subject)
```

sjekk at full_treatment blir registert

```{r}
# denne måtte skilles ut for å bli registrert riktig!?
data_emne2_v23$full_treatment[data_emne2_v23$Subject %in% data_post_emne2$Subject] <- 1

table(data_emne2_v23$full_treatment)
```

```{r create_dataframe_v22}

# make working copy
data_emne2_v22    <-  data_TPACK_pilot  [data_TPACK_pilot$datasett == "pre" | 
                                           data_TPACK_pilot$datasett == "post",
                                         !colnames(data_TPACK_pilot) %in%
                                           c("semester","studiested",
                                             "XK1","XK2","XK3","XK4","XK5","XK6","CK123","CK6")]

```

Sort datasets

```{r sort_students}
# sort students
# sorter på person
data_emne2_v22$datasett <- factor(data_emne2_v22$datasett, level=c("pre","post"))
data_emne2_v22 <-  data_emne2_v22[
                          with(data_emne2_v22, 
                          order(Subject,datasett, # rekkefølge
                          method="radix", # for faktorer, brukes med c nedenfor
                          decreasing = c(0,0))),] # bare datasett (1)

# set correct order of dataset
data_emne2_v23$datasett <- factor(data_emne2_v23$datasett, level=c("pre1","pre2","post"))
data_emne2_v23 <-  data_emne2_v23[
                          with(data_emne2_v23, 
                          order(Subject,datasett, # rekkefølge
                          method="radix", # for faktorer, brukes med c nedenfor
                          decreasing = c(0,0))),] # bare datasett (1) 
```

## Instruments (data collection)

Data were collected using the measure developed by @karlsen_assessing_2024 (paper 1).

The measure used in paper 4 was an early version of the instrument developed in paper 1. This led to the TK-items being somewhat different from the ones in paper 1. Here I use TK2, TK5 and TK6, while in paper 1 I use TK1, TK2 and TK3.

## Analyses

### Choose subsample of PSTs

For easier comparison between the different analysis methods, the same sample is used for the analyses (i.e., n=19 for prototype1 and n=13 for prototype 2).

Select students

```{r select_students_v22}
###
# v22
###

# bare se på studenter som er med både pre og post
stud_alle_v22  <- intersect(data_emne2_v22$Subject[data_emne2_v22$datasett == "pre"],
                         data_emne2_v22$Subject[data_emne2_v22$datasett == "post"])

# hvilke studenter var ikke med på pre, men på post? 
stud_bare_pre_eller_post_v22  <- data_emne2_v22$Subject[!data_emne2_v22$Subject %in%
                                                          stud_alle_v22]

# hvilke var med bare på post?
stud_bare_post_v22 <- data_emne2_v22$Subject[
  data_emne2_v22$datasett == "post" &
    data_emne2_v22$Subject %in% stud_bare_pre_eller_post_v22]

data_emne2_v22$testing <- rep(0,nrow(data_emne2_v22))
data_emne2_v22$testing[data_emne2_v22$Subject == stud_bare_post_v22] <- 1

alle_studenter_v22 <- data_emne2_v22$Subject %in% stud_alle_v22
```

```{r select_students_v23} 
###
# v23
###

# bare se på studenter som er med både pre og post
stud_alle_v23  <- intersect(data_emne2_v23$Subject[data_emne2_v23$datasett == "pre1"],
                         data_emne2_v23$Subject[data_emne2_v23$datasett == "pre2"])

# hvilke var med bare på pre2?
stud_bare_pre2_v23 <- data_emne2_v23$Subject[
  data_emne2_v23$datasett == "pre2" &
    !data_emne2_v23$Subject %in% stud_alle_v23]

stud_alle_v23  <- intersect(stud_alle_v23,
                         data_emne2_v23$Subject[data_emne2_v23$datasett == "post"])

# hvilke var med bare på post?
stud_bare_post_v23 <- data_emne2_v23$Subject[
  data_emne2_v23$datasett == "post" &
    !data_emne2_v23$Subject %in% stud_alle_v23]

data_emne2_v23$testing <- rep(0,nrow(data_emne2_v23))
data_emne2_v23$testing[data_emne2_v23$Subject %in% stud_bare_pre2_v23] <- 1

#data_emne2_v22 <- data_emne2_v22[data_emne2_v22$Subject %in% stud_alle_v22,]
#data_emne2_v23 <- data_emne2_v23[data_emne2_v23$Subject %in% stud_alle_v23,]

alle_studenter_v23 <- data_emne2_v23$Subject %in% stud_alle_v23
```

### Difference scores

To decide which statistical tests we should use to compare pre-post scores, the Levene’s (Mean) test for equal variances and the Kolmogorov-Smirnov test for normality were used as preliminary tests [@pearce_preliminary_2019]. Due to the small sample sizes, the data are probably non-normally distributed. If the variance of the pre-post differences are unequal, this suggests the use of the non-parametric Yuen d-test with trimmed means [@mair_robust_2020], which show good power for low samples [@wanishsakpong_comparison_2023] and can be appropriate when one is not interested in generalizing to a population [@ng_gamma_2019]. If, on the other hand, the pre-post data have equal variance, this suggests the use of the Mann-Whitney (i.e., Wilcox) test with arithmetic mean.

Hedges' g was also calculated, since it is often used in meta-analyses.

#### Prototype 1

```{r descriptive_stat_v22}
scores_v22 <- scoreItems(TPACK_PST_emne2_v22, data_emne2_v22)
data_emne2_v22$TK <- scores_v22$scores[,"TK"]
data_emne2_v22$PCK <- scores_v22$scores[,"PCK"]
data_emne2_v22$TPACK <- scores_v22$scores[,"TPACK"]

#describe(data_emne2_v22[data_emne2_v22$datasett == "pre",])
scores_v22_pre <- scoreItems(TPACK_PST_emne2_v22,
                             data_emne2_v22[data_emne2_v22$datasett == "pre" &
                                            data_emne2_v22$Subject %in%
                                              stud_alle_v22,])
describe(scores_v22_pre$scores)

#describe(data_emne2_v22[data_emne2_v22$datasett == "post",])
scores_v22_post <- scoreItems(TPACK_PST_emne2_v22,
                              data_emne2_v22[data_emne2_v22$datasett == "post" &
                                             data_emne2_v22$Subject %in%
                                               stud_alle_v22,])
describe(scores_v22_post$scores)
```

```{r create_table_v22}
#!NB Sjekk om rekkefølgen her er riktig? Eller alternativt bruk data_emne2_v22-dataframen

# gjør om til loop

# pre
TK_pre <- scores_v22_pre$scores[,"TK"]
TK_alpha_pre <- scores_v22_pre$corrected["TK","TK"]

PCK_pre <- scores_v22_pre$scores[,"PCK"]
PCK_alpha_pre <- scores_v22_pre$corrected["PCK","PCK"]

TPACK_pre <- scores_v22_pre$scores[,"TPACK"]
TPACK_alpha_pre <- scores_v22_pre$corrected["TPACK","TPACK"]

#post
TK_post <- scores_v22_post$scores[,"TK"]
TK_alpha_post <- scores_v22_post$corrected["TK","TK"]

PCK_post <- scores_v22_post$scores[,"PCK"]
PCK_alpha_post <- scores_v22_post$corrected["PCK","PCK"]

TPACK_post <- scores_v22_post$scores[,"TPACK"]
TPACK_alpha_post <- scores_v22_post$corrected["TPACK","TPACK"]

# cohens d
TK_cd <- effsize::cohen.d(TK_post, TK_pre, 
                 hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
PCK_cd <- effsize::cohen.d(PCK_post, PCK_pre, 
                  hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
TPACK_cd <- effsize::cohen.d(TPACK_post, TPACK_pre, 
                    hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)

# tabell
TK_rad <- c(mean(TK_pre), mean(TK_post), TK_cd$estimate, TK_cd$conf.int, TK_alpha_pre, TK_alpha_post)
PCK_rad <- c(mean(PCK_pre), mean(PCK_post), PCK_cd$estimate, PCK_cd$conf.int, PCK_alpha_pre, PCK_alpha_post)
TPACK_rad <- c(mean(TPACK_pre), mean(TPACK_post), TPACK_cd$estimate, TPACK_cd$conf.int, TPACK_alpha_pre, TPACK_alpha_post)
tabell_v22 <- t(data.frame(TK_rad, #TCK_rad, TPK_rad, 
                           PCK_rad, TPACK_rad))
rownames(tabell_v22) <- c("TK", "PCK", "TPACK")
colnames(tabell_v22) <- c("pre","post", "cohen's d", "ci lower", "ci upper", "alpha_pre", "alpha_post")

pander(tabell_v22)
```

Er det noen signifikant forskjell mellom de ulike periodene?

sjekk forutsetninger for t-test (normalitet og lik varians)

```{r}
# sjekk av antagelser
# ref. Pearce and Derrick (2019)

TK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "TK"]
TK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "TK"]
PCK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "PCK"]
PCK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "PCK"]
TPACK22_pre <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "pre", "TPACK"]
TPACK22_post <- data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett == "post", "TPACK"]

df22_prepost <- data.frame(TK = c(TK22_pre, TK22_post),
                    PCK = c(PCK22_pre, PCK22_post),
                    TPACK = c(TPACK22_pre, TPACK22_post),
                    prototype = c(rep("pre", length(TK22_pre)), 
                                  rep("post", length(TK22_post))))

```

```{r}
# Levene's (Mean) test for equal variances
# ref https://www.statology.org/levenes-test-r/
#library(car)
leveneTest(TK ~ prototype, center=mean, data = df22_prepost)
leveneTest(PCK ~ prototype, center=mean, data = df22_prepost)
leveneTest(TPACK ~ prototype, center=mean, data = df22_prepost)
```

Ulik varians.

```{r}
ks.test(TK22_pre, "pnorm")
ks.test(TK22_post, "pnorm")
ks.test(PCK22_pre, "pnorm")
ks.test(PCK22_post, "pnorm")
ks.test(TPACK22_pre, "pnorm")
ks.test(TPACK22_post, "pnorm")
```

Ikke normalfordelt.

```{r}
# Yuen-Welch test
# ref.: Mair & Wilcox, 2020
#library(WRS2)
yuend(TK22_pre, TK22_post)
yuend(PCK22_pre, PCK22_post)
yuend(TPACK22_pre, TPACK22_post)
```

```{r}
t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "TK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "TK"], 
       paired=T)

t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "PCK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "PCK"], 
       paired=T)

t.test(data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="pre", "TPACK"],
       data_emne2_v22[alle_studenter_v22 & data_emne2_v22$datasett=="post", "TPACK"], 
       paired=T)
```

Se på forandring i TPACK

```{r calc_diff_v22}
calc_diff_TPACK_v22 <- function (TPACK_element) {

df <- data_emne2_v22[data_emne2_v22$datasett == "pre" | 
                     data_emne2_v22$datasett == "post",
                     c("Subject","datasett",TPACK_element)] # "group",

# sorter på datasett
df <-  df[
                          with(df, 
                          order(datasett,Subject, # rekkefølge
                          method="radix"#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(-1,0))
                          ))
                          ,] # bare datasett (1)

# regn ut TPACK pre/post for hver person
ifelse (length(TPACK_element) > 1, {
TPACK_pre <- rowMeans(df[df$datasett=="pre",3:ncol(df)],na.rm=TRUE)
TPACK_post <- rowMeans(df[df$datasett=="post",3:ncol(df)],na.rm=TRUE)
},
{
  TPACK_pre <- df[df$datasett=="pre",TPACK_element]
  TPACK_post <- df[df$datasett=="post",TPACK_element]
})
TPACK_merge <- c(TPACK_pre,TPACK_post)

# legg til resultat
df$TPACK_pre_post <- TPACK_merge

# sorter på person; 
df <-  df[
                          with(df, 
                          order(Subject,datasett, # rekkefølge
                          method="radix"))#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(0,1)))
                          ,] # bare datasett (1)

# behold bare personer som er med i  både pre/post
df_filtered <- df[!df$Subject %in% c("Benjamin","Gabriel","Olava"),]

# beregn differanse; pre inneholder riktig diff-verdi
df_filtered$diff <-
  c(diff(df_filtered[,ncol(df_filtered)]),0)

diff_TPACK <- df_filtered[df_filtered$datasett=="pre",c(1,ncol(df_filtered))]

return (diff_TPACK)
}

TK_diff_v22 <- calc_diff_TPACK_v22(TK_PST22_pos)
PCK_diff_v22 <- calc_diff_TPACK_v22(PCK_PST22_pos)
TPACK_diff_v22 <- calc_diff_TPACK_v22(TPACK_PST22_pos)

# lag dataframe
diff_full_v22 <- data.frame("Subject"=TK_diff_v22[,1], #"group" = TK_diff[,"group"],
  "TK"=TK_diff_v22[,2],
  "PCK"=PCK_diff_v22[,2],
  "TPACK"=TPACK_diff_v22[,2])
```

##### reliable change in individual scores

beregn SID [@estrada_statistics_2019].

```{r}
SID <- TK_diff_v22[,2] / SD(TK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- PCK_diff_v22[,2] / SD(PCK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- TPACK_diff_v22[,2] / SD(TPACK_diff_v22[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

##### Sjekk for RTM

RTM = regression to the mean

Ref.: @marsden_single_2012

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"TK"], 
                  "TK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"TK"]),
       aes(x=TK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"PCK"], 
                  "PCK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"PCK"]),
       aes(x=PCK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v22[,"TPACK"], 
                  "TPACK" = data_emne2_v22[data_emne2_v22$datasett == "pre" & 
                                        data_emne2_v22$Subject %in% stud_alle_v22,"TPACK"]),
       aes(x=TPACK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

#### Prototype 2

```{r descriptive_stat_v23}
scores_v23 <- scoreItems(TPACK_PST_emne2_v23, data_emne2_v23)
data_emne2_v23$TK <- scores_v23$scores[,"TK"]
data_emne2_v23$PCK <- scores_v23$scores[,"PCK"]
data_emne2_v23$TPACK <- scores_v23$scores[,"TPACK"]

#describe(data_emne2_v23[data_emne2_v23$datasett == "pre",])
scores_v23_pre <- scoreItems(TPACK_PST_emne2_v23, data_emne2_v23[data_emne2_v23$datasett == "pre1" & data_emne2_v23$Subject %in% stud_alle_v23,])
describe(scores_v23_pre$scores)

#describe(data_emne2_v23[data_emne2_v23$datasett == "post",])
scores_v23_post <- scoreItems(TPACK_PST_emne2_v23,
                              data_emne2_v23[data_emne2_v23$datasett == "post" &
                                             data_emne2_v23$Subject %in%
                                               stud_alle_v23,])
describe(scores_v23_post$scores)
```

```{r create_table_v23}
#!NB Sjekk om rekkefølgen her er riktig? Eller alternativt bruk data_emne2_v23-dataframen

# gjør om til loop

# pre
TK_pre <- scores_v23_pre$scores[,"TK"]
TK_alpha_pre <- scores_v23_pre$corrected["TK","TK"]

PCK_pre <- scores_v23_pre$scores[,"PCK"]
PCK_alpha_pre <- scores_v23_pre$corrected["PCK","PCK"]

TPACK_pre <- scores_v23_pre$scores[,"TPACK"]
TPACK_alpha_pre <- scores_v23_pre$corrected["TPACK","TPACK"]

#post
TK_post <- scores_v23_post$scores[,"TK"]
TK_alpha_post <- scores_v23_post$corrected["TK","TK"]

PCK_post <- scores_v23_post$scores[,"PCK"]
PCK_alpha_post <- scores_v23_post$corrected["PCK","PCK"]

TPACK_post <- scores_v23_post$scores[,"TPACK"]
TPACK_alpha_post <- scores_v23_post$corrected["TPACK","TPACK"]

# cohens d
TK_cd <- effsize::cohen.d(TK_post, TK_pre, 
                 hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
PCK_cd <- effsize::cohen.d(PCK_post, PCK_pre, 
                  hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)
TPACK_cd <- effsize::cohen.d(TPACK_post, TPACK_pre, 
                    hedges.correction=TRUE, within=T, paired=T, noncentral=T)#, conf.level=0.80)

# tabell
TK_rad <- c(mean(TK_pre), mean(TK_post), TK_cd$estimate, TK_cd$conf.int, TK_alpha_pre, TK_alpha_post)
PCK_rad <- c(mean(PCK_pre), mean(PCK_post), PCK_cd$estimate, PCK_cd$conf.int, PCK_alpha_pre, PCK_alpha_post)
TPACK_rad <- c(mean(TPACK_pre), mean(TPACK_post), TPACK_cd$estimate, TPACK_cd$conf.int, TPACK_alpha_pre, TPACK_alpha_post)
tabell_v23 <- t(data.frame(TK_rad, #TCK_rad, TPK_rad, 
                           PCK_rad, TPACK_rad))
rownames(tabell_v23) <- c("TK", "PCK", "TPACK")
colnames(tabell_v23) <- c("pre","post", "cohen's d", "ci lower", "ci upper", "alpha_pre", "alpha_post")

pander(tabell_v23)
```

Er det noen signifikant forskjell mellom de ulike periodene?

sjekk forutsetninger for t-test (normalitet og lik varians)

```{r}
# sjekk av antagelser
# ref. Pearce and Derrick (2019)

TK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "TK"]
TK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "TK"]
PCK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "PCK"]
PCK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "PCK"]
TPACK23_pre <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "pre1", "TPACK"]
TPACK23_post <- data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett == "post", "TPACK"]

df23_prepost <- data.frame(TK = c(TK23_pre, TK23_post),
                    PCK = c(PCK23_pre, PCK23_post),
                    TPACK = c(TPACK23_pre, TPACK23_post),
                    prototype = c(rep("pre1", length(TK23_pre)), 
                                  rep("post", length(TK23_post))))

```

```{r}
# Levene's (Mean) test for equal variances
# ref https://www.statology.org/levenes-test-r/
#library(car)
leveneTest(TK ~ prototype, center=mean, data = df23_prepost)
leveneTest(PCK ~ prototype, center=mean, data = df23_prepost)
leveneTest(TPACK ~ prototype, center=mean, data = df23_prepost)
```

Ulik varians.

```{r}
ks.test(TK23_pre, "pnorm")
ks.test(TK23_post, "pnorm")
ks.test(PCK23_pre, "pnorm")
ks.test(PCK23_post, "pnorm")
ks.test(TPACK23_pre, "pnorm")
ks.test(TPACK23_post, "pnorm")
```

Ikke normalfordelt.

```{r}
# Wilcoxon signed-rank test (Mann-Whitney)
wilcox.test(TK23_pre, TK23_post, paired=T)
wilcox.test(PCK23_pre, PCK23_post, paired=T)
wilcox.test(TPACK23_pre, TPACK23_post, paired=T)
```

```{r}
# Yuen-Welch test
# ref.: Mair & Wilcox, 2020
#library(WRS2)
yuend(TK23_pre, TK23_post)
yuend(PCK23_pre, PCK23_post)
yuend(TPACK23_pre, TPACK23_post)
```

```{r}
t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "TK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "TK"], 
       paired=T)

t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "PCK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "PCK"], 
       paired=T)

t.test(data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="pre1", "TPACK"],
       data_emne2_v23[alle_studenter_v23 & data_emne2_v23$datasett=="post", "TPACK"], 
       paired=T)
```

Se på forandring i TPACK

```{r calc_diff_v23}
calc_diff_TPACK_v23 <- function (TPACK_element) {

df <- data_emne2_v23[data_emne2_v23$datasett == "pre1" | 
                     data_emne2_v23$datasett == "post",
                     c("Subject","datasett",TPACK_element)] # "group",

# sorter på datasett
df <-  df[
                          with(df, 
                          order(datasett,Subject, # rekkefølge
                          method="radix"#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(-1,0))
                          ))
                          ,] # bare datasett (1)

# regn ut TPACK pre/post for hver person
ifelse (length(TPACK_element) > 1, {
TPACK_pre <- rowMeans(df[df$datasett=="pre1",3:ncol(df)],na.rm=TRUE)
TPACK_post <- rowMeans(df[df$datasett=="post",3:ncol(df)],na.rm=TRUE)
},
{
  TPACK_pre <- df[df$datasett=="pre1",TPACK_element]
  TPACK_post <- df[df$datasett=="post",TPACK_element]
})
TPACK_merge <- c(TPACK_pre,TPACK_post)

# legg til resultat
df$TPACK_pre_post <- TPACK_merge

# sorter på person; 
df <-  df[
                          with(df, 
                          order(Subject,datasett, # rekkefølge
                          method="radix"))#, # for faktorer, brukes med c nedenfor
                          #decreasing = c(0,1)))
                          ,] # bare datasett (1)

# behold bare personer som er med i  både pre/post
df_filtered <- df[df$Subject %in% stud_alle_v23,]

# beregn differanse; pre inneholder riktig diff-verdi
df_filtered$diff <-
  c(diff(df_filtered[,ncol(df_filtered)]),0)

diff_TPACK <- df_filtered[df_filtered$datasett=="pre1",c(1,ncol(df_filtered))]

return (diff_TPACK)
}

TK_diff_v23 <- calc_diff_TPACK_v23(TK_PST23_pos)
PCK_diff_v23 <- calc_diff_TPACK_v23(PCK_PST23_pos)
TPACK_diff_v23 <- calc_diff_TPACK_v23(TPACK_PST23_pos)

# lag dataframe
diff_full_v23 <- data.frame("Subject"=TK_diff_v23[,1], #"group" = TK_diff[,"group"],
  "TK"=TK_diff_v23[,2],
  "PCK"=PCK_diff_v23[,2],
  "TPACK"=TPACK_diff_v23[,2])
```

##### reliable change in individual scores

beregn SID [@estrada_statistics_2019].

```{r}
SID <- TK_diff_v23[,2] / SD(TK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- PCK_diff_v23[,2] / SD(PCK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

```{r}
SID <- TPACK_diff_v23[,2] / SD(TPACK_diff_v23[,2])

# percentage of reliable improvements
pri <- sum(SID > 1.645, na.rm=TRUE)
pri
pri / length(SID)
```

##### Sjekk for RTM

RTM = regression to the mean

Ref.: @marsden_single_2012

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"TK"], 
                  "TK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"TK"]),
       aes(x=TK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"PCK"], 
                  "PCK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"PCK"]),
       aes(x=PCK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

```{r}
#plot diff vs pre
ggplot(data.frame("postpre" = diff_full_v23[,"TPACK"], 
                  "TPACK" = data_emne2_v23[data_emne2_v23$datasett == "pre1" & 
                                        data_emne2_v23$Subject %in% stud_alle_v23,"TPACK"]),
       aes(x=TPACK,y=postpre)) +
  geom_jitter()  +
  stat_smooth(method = "lm", formula = 'y ~ x', se=F,fullrange = F)
```

### Fixed-effects

Before doing the regression analysis, we need to decide how to model the data. Should we model it through pooled OLS, fixed effects or random effects?

... Legg til sjekker av hvilken modell som bør brukes

```{r}
within22 <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v22,
              index=c("Subject","datasett"), model="within")
firstdifference22 <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v22,
              index=c("Subject","datasett"), model="fd")
```

```{r within_3pt}
within23 <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v23,
              index=c("Subject","datasett"), model="within")
firstdifference23 <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v23,
              index=c("Subject","datasett"), model="fd")
```

```{r within_2pt}
within23_prepost <- plm(TPACK ~ TK + PCK + factor(Subject) + factor(datasett), 
              data=data_emne2_v23[data_emne2_v23$datasett == "pre1" | data_emne2_v23$datasett == "post",],
              index=c("Subject","datasett"), model="within")
firstdifference23_prepost <- plm(TPACK ~ TK + PCK + factor(Subject),# + factor(datasett),
              data=data_emne2_v23[data_emne2_v23$datasett == "pre1" | data_emne2_v23$datasett == "post",],
              index=c("Subject","datasett"), model="fd")
```

```{r}
screenreg(list("within22" = within22,
               "fd22" = firstdifference22,
               "within23" = within23,
               "fd23" = firstdifference23,
               "within23pp" = within23_prepost,
               "fd23pp" = firstdifference23_prepost))
```

Fixed effects were used, comparing only pre-post.

### Within-between models

To calculate the within-between model for prototype 2, all three datapoints were needed. This was not done in paper 4.

Prototype 1

prepare centred and demeaned variabels

create a dummy variable for each individual
or
gruppesentrering, _S

```{r prot1_centering}
# TK
# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
data_emne2_v22$TK_mean <- with(data_emne2_v22, ave(x = TK, Subject, FUN = mean))
data_emne2_v22$TK_S <- data_emne2_v22$TK - data_emne2_v22$TK_mean

# PCK interval
data_emne2_v22$PCK_mean <- with(data_emne2_v22, ave(x = PCK, Subject, FUN = mean))
data_emne2_v22$PCK_S <- data_emne2_v22$PCK - data_emne2_v22$PCK_mean

# TPACK
data_emne2_v22$TPACK_mean <- with(data_emne2_v22, ave(x = TPACK, Subject, FUN = mean))
data_emne2_v22$TPACK_S <- data_emne2_v22$TPACK - data_emne2_v22$TPACK_mean
```

```{r prot2_centering}
# TK
# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
data_emne2_v23$TK_mean <- with(data_emne2_v23, ave(x = TK, Subject, FUN = mean))
data_emne2_v23$TK_S <- data_emne2_v23$TK - data_emne2_v23$TK_mean

# PCK interval
data_emne2_v23$PCK_mean <- with(data_emne2_v23, ave(x = PCK, Subject, FUN = mean))
data_emne2_v23$PCK_S <- data_emne2_v23$PCK - data_emne2_v23$PCK_mean

# TPACK
data_emne2_v23$TPACK_mean <- with(data_emne2_v23, ave(x = TPACK, Subject, FUN = mean))
data_emne2_v23$TPACK_S <- data_emne2_v23$TPACK - data_emne2_v23$TPACK_mean
```

```{r prot2_centering_prepost}
# do not include pre2-data
data_emne2_v23_pp <- data_emne2_v23[!data_emne2_v23$datasett == "pre2",]

alle_studenter_v23_pp <-
  alle_studenter_v23[!data_emne2_v23$datasett == "pre2"]

# TK
# calculate mean for each subject / id; ref.: https://stackoverflow.com/a/34502529
data_emne2_v23_pp$TK_mean <- with(data_emne2_v23_pp, ave(x = TK, Subject, FUN = mean))
data_emne2_v23_pp$TK_S <- data_emne2_v23_pp$TK - data_emne2_v23_pp$TK_mean

# PCK interval
data_emne2_v23_pp$PCK_mean <- with(data_emne2_v23_pp, ave(x = PCK, Subject, FUN = mean))
data_emne2_v23_pp$PCK_S <- data_emne2_v23_pp$PCK - data_emne2_v23_pp$PCK_mean

# TPACK
data_emne2_v23_pp$TPACK_mean <- with(data_emne2_v23_pp, ave(x = TPACK, Subject, FUN = mean))
data_emne2_v23_pp$TPACK_S <- data_emne2_v23_pp$TPACK - data_emne2_v23_pp$TPACK_mean
```

Hybrid with variable coefficients was chosen since there was only one significant dependent variable in the fixed-effects model [@bell_fixed_2019].

Ville det vært bedre å sammenligne med en enkel hybrid model uten variable koeffisienter først, for mer direkte sammenligning med FE-modellen?

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TPACK_n19 <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

summary(hybrid_TPACK_n19)
```

bare pre-post

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TPACK_varcof <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

summary(hybrid_TPACK_varcof)
```

alle studentene

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

hybrid_TPACK <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22)

summary(hybrid_TPACK)
```

#### Comparison hybrid models

Use FIML (ML) for model comparison, REML for parameter estimates [@christophersen_introduksjon_2018].

Prototype 1

```{r eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

prot1_within_between <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="ML", data=data_emne2_v22[alle_studenter_v22,])

prot1_var_coeff <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="ML", data=data_emne2_v22[alle_studenter_v22,])
```

Prototype 2

```{r prot2_wb, eval=T}
# TK_S within effects
# TK_mean between effects
# factor(datasett) fixed time effects
# TK_S|id variable coefficient TK_S per individual

# only 13 from both pre-post
prot2_within_between_13 <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", 
                   data=data_emne2_v23_pp[alle_studenter_v23_pp,])

# all 30 for comparison with var coeff below
prot2_within_between <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # [alle_studenter_v23,]
```

using data_emne2_v23[ !data_emne2_v23$datasett == "pre2",] gives an error in the estimate, since TK_mean, etc. are calculated using the mean for all three datapoints, not only pre1 and post. It does not converge if I use the _pp dataset.

```{r prot2_vc}
# does not converge with only 13 respondents
prot2_var_coeff <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="ML", data=data_emne2_v23[ !data_emne2_v23$datasett == "pre2",]) # 
```

```{r}
# print summary of models
screenreg(list("P1_wb" = prot1_within_between, 
               "P1_vc" = prot1_var_coeff,
               "P2_wb" = prot2_within_between, 
               "P2_vc" = prot2_var_coeff,
               "P2_13" = prot2_within_between_13))
```

chi-square comparison, significant difference between models?

```{r}
anova(prot1_within_between, prot1_var_coeff)

anova(prot2_within_between, prot2_var_coeff)
```

AICc for small samples

```{r}
AICc(prot1_within_between, prot1_var_coeff, REML=F)
AICc(prot2_within_between, prot2_var_coeff, REML=F)
```

#### Summary of details for best fit models

```{r}
prot1_var_coeff_REML <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~TK_S|Subject, method="REML", data=data_emne2_v22[alle_studenter_v22,])

prot2_within_between_13_REML <- lme(TPACK ~ TK_S + TK_mean + PCK_S + PCK_mean + factor(datasett), 
                   random=~1|Subject, method="REML", 
                   data=data_emne2_v23_pp[alle_studenter_v23_pp,])

summary(prot1_var_coeff_REML)
summary(prot2_within_between_13_REML)
```

# Results

Here I reproduce the results presented in paper 4.

## Difference

```{r}
# v22
pre_scores_v22 <- describe(scores_v22_pre$scores)
post_scores_v22 <- describe(scores_v22_post$scores)

diff_TK_v22 <- yuend(scores_v22_post$scores[,"TK"],
                  scores_v22_pre$scores[,"TK"]
                  )
diff_PCK_v22 <- yuend(scores_v22_post$scores[,"PCK"],
                  scores_v22_pre$scores[,"PCK"]
                  )
diff_TPACK_v22 <- yuend(scores_v22_post$scores[,"TPACK"],
                  scores_v22_pre$scores[,"TPACK"]
                  )

# v23
pre_scores_v23 <- describe(scores_v23_pre$scores)
post_scores_v23 <- describe(scores_v23_post$scores)

diff_TK_v23 <- post_scores_v23["TK","mean"] - pre_scores_v23["TK","mean"]
diff_p_TK_v23 <- wilcox.test(scores_v23_post$scores[,"TK"], scores_v23_pre$scores[,"TK"], paired=T)$p.value

diff_PCK_v23 <- post_scores_v23["PCK","mean"] - pre_scores_v23["PCK","mean"]
diff_p_PCK_v23 <- wilcox.test(scores_v23_post$scores[,"PCK"], scores_v23_pre$scores[,"PCK"], paired=T)$p.value

diff_TPACK_v23 <- post_scores_v23["TPACK","mean"] - pre_scores_v23["TPACK","mean"]
diff_p_TPACK_v23 <- wilcox.test(scores_v23_post$scores[,"TPACK"], scores_v23_pre$scores[,"TPACK"], paired=T)$p.value

# create table
rad_TK_v22 <- list("TK",
  pre_scores_v22["TK","mean"],
            pre_scores_v22["TK","sd"],
            post_scores_v22["TK","mean"],
            post_scores_v22["TK","sd"],
            paste0(format(diff_TK_v22$diff,digit=2),
                   " (p = ", 
                   format(diff_TK_v22$p.value, digits=2), 
                   ")"))

rad_PCK_v22 <- list("PCK",
  pre_scores_v22["PCK","mean"],
            pre_scores_v22["PCK","sd"],
            post_scores_v22["PCK","mean"],
            post_scores_v22["PCK","sd"],
            paste0(format(diff_PCK_v22$diff,digit=2),
                   " (p = ", 
                   format(diff_PCK_v22$p.value, digits=2), 
                   ")"))

rad_TPACK_v22 <- list("TPACK",
  pre_scores_v22["TPACK","mean"],
            pre_scores_v22["TPACK","sd"],
            post_scores_v22["TPACK","mean"],
            post_scores_v22["TPACK","sd"],
            paste0(format(diff_TPACK_v22$diff,digit=2),
                   " (p = ", 
                   format(diff_TPACK_v22$p.value, digits=2), 
                   ")"))

rad_TK_v23 <- list("TK",
            pre_scores_v23["TK","mean"],
            pre_scores_v23["TK","sd"],
            post_scores_v23["TK","mean"],
            post_scores_v23["TK","sd"],
            paste0(format(diff_TK_v23, digit=2),
                   " (p = ",
                   format(diff_p_TK_v23, digit=2),
                   ")")
            )

rad_PCK_v23 <- list("PCK",
            pre_scores_v23["PCK","mean"],
            pre_scores_v23["PCK","sd"],
            post_scores_v23["PCK","mean"],
            post_scores_v23["PCK","sd"],
            paste0(format(diff_PCK_v23, digit=2),
                   " (p = ",
                   format(diff_p_PCK_v23, digit=2),
                   ")")
            )

rad_TPACK_v23 <- list("TPACK",
            pre_scores_v23["TPACK","mean"],
            pre_scores_v23["TPACK","sd"],
            post_scores_v23["TPACK","mean"],
            post_scores_v23["TPACK","sd"],
            paste0(format(diff_TPACK_v23, digit=2, nsmall=2),
                   " (p = ",
                   format(diff_p_TPACK_v23, digit=2, nsmall=2),
                   ")")
            )

row1_p1 <- c("", "Prototype 1", "", "", "", "")
row1_p2 <- c("", "Prototype 2", "", "", "", "")
row2 <- c("", "Pre", "", "Post", "", "")
row3 <- c("", "Mean", "SD", "Mean", "SD", "diff")

prepost_table_v22 <- t(data.frame(
  row1_p1, row2, row3,
  format(rad_TK_v22,digit=2,nsmall=2), # only 2 digits
  format(rad_PCK_v22,digit=2,nsmall=2), # only 2 digits
  format(rad_TPACK_v22,digit=2,nsmall=2) # only 2 digits
))

prepost_table_v23 <- t(data.frame(
  row1_p2, row2, row3,
  format(rad_TK_v23,digit=2,nsmall=2), # only 2 digits
  format(rad_PCK_v23,digit=2,nsmall=2), # only 2 digits
  format(rad_TPACK_v23,digit=2,nsmall=2) # only 2 digits
))

rownames(prepost_table_v22) <- NULL # remove row names
rownames(prepost_table_v23) <- NULL # remove row names

pander(prepost_table_v22)
pander(prepost_table_v23)
```

## Fixed effects

```{r}
screenreg(list("within22" = within22,
               "within23pp" = within23_prepost))
```

## Within-between

```{r}
summary(hybrid_TPACK)
```

TK_S and PCK_S are the within-effects, while TK_mean and PCK_mean are the between-effects.

This model is however not directly comparably to the fixed effects model, since the hybrid model uses all the students in the analysis (n=21+20), while the fixed effects only uses the students that are in both pre-post datasets (n=19+19) (cf. model hybrid_TPACK_n19 above).

```{r}
# print summary of models
screenreg(list("Hybrid (all students)" = hybrid_TPACK, 
               "Hybrid (pre-post students)" = hybrid_TPACK_n19))
```

# Discussion

The hybrid model reported in paper 4, should perhaps be switched to the n=19 model, for more direct comparison with the fixed effects model.

# References

<!-- citation(package) -->

<div id="refs"></div>

<!-- source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html

# (APPENDIX) Appendix {-} -->

# Session info
<!-- # ref.: https://intro2r.com/proj_doc.html -->
<!-- #sessionInfo() -->

```{r}
xfun::session_info()
```